---
title: "Mouse Hierarchy"
author: "Rinaldo D'Souza"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# install beta regression package; click through the dialogue boxes that appear

```{r}
if(!require(betareg)) install.packages("betareg") else
  suppressPackageStartupMessages(library(betareg))
```


# enter data 

```{r}
library(readxl)
connectivity_matrix <- read_excel("C:\\Users\\rinal\\OneDrive\\Desktop\\Rinaldo (Desktop)\\mouse hierarchy analysis\\R files\\NewExcel\\Average_L24toL14_70saturation_blur1um.xlsx")
m <- connectivity_matrix
m <- m[, -1]
m <- as.matrix(m)
rownames(m) <- colnames(m)
#round(m, 2)
m
```

# p.11: set up incidence matrix for testing models (for beta regression, etc.)

```{r}
m.df <- stack(as.data.frame(t(m)))
names(m.df) <- c("ODr", "Source")
m.df$Target <- factor(colnames(m), levels = colnames(m))
m.df$Source <- factor(m.df$Source, levels = colnames(m))
Imat <- matrix(0, nrow = 100, ncol = 10)
colnames(Imat) <- colnames(m)
rownames(Imat) <- with(m.df, 
                       paste0(as.character(Source), "->", as.character(Target)))
for(ix in 1:100){
  Imat[ix, m.df[ix, "Source"]] <- -1
  Imat[ix, m.df[ix, "Target"]] <- 1
  }
Imat <- Imat[complete.cases(m.df), ]    
m.df <- m.df[complete.cases(m.df), ]
Dorsal <- c("LM", "P", "LI", "POR")
Ventral <- c("AL", "RL", "A", "PM", "AM")
m.df$Stream <- with(m.df, 
                ifelse(Source %in% Dorsal, "D",
                  ifelse(Source %in% Ventral, "V", "V1")))
```


# p.3: print results of binomial test
```{r}
binom.test(27, 37)
```

binom.test(27, 37)

# p.4: check for symmetry of FF and FB strengths

```{r}
linfffb.lm <- lm(upm ~ lom)
summary(linfffb.lm)
```

```{r}
fffb.lm <- lm(qlogis(upm) ~ qlogis(lom))
summary(fffb.lm)
```
# p.5: plot FF vs FB values

```{r, fig.align = "center"}
par(pty = "s")
plot(upm ~ lom, xlab  ="DR lower matrix",
     ylab = "DR upper matrix")
xx <- seq(-0.5, 2, len = 100)
linfffb.pr <- predict(linfffb.lm, newdata = data.frame(lom = xx), 
                   se.fit = TRUE, interval = "prediction")
lines(xx, linfffb.pr$fit[, 2], lty = 2)
lines(xx, linfffb.pr$fit[, 3], lty = 2)
abline(linfffb.lm)
```


# p.7: triangle plot: check OD ratios for each area against unit slope line

```{r, fig.width = 7, fig.height = 7, fig.align = "center"}
pairs(t(m), xlim = c(0, 0.8), ylim = c(0, 0.8),
	lower.panel = function(x, y, ...){
	  points(x, y)
		xy0 <- lm(y ~ offset(x))
		abline(coef(xy0), 1, lty = 2, lwd = 2)
	}, upper.panel = NULL, gap = 0.2)
```

```{r, fig.width = 7, fig.height = 7, fig.align = "center"}
pairs(qlogis(m), xlim = c(-2, 2), ylim = c(-2, 2),
	lower.panel = function(x, y, ...){
	  points(x, y)
		xy0 <- lm(y ~ offset(x))
		abline(coef(xy0), 1, lty = 2, lwd = 2)
	}, upper.panel = NULL, gap = 0.2)
```

```{r}
m.df <- stack(as.data.frame(t(m)))
names(m.df) <- c("ODr", "Source")
m.df$Target <- factor(colnames(m), levels = colnames(m))
m.df$Source <- factor(m.df$Source, levels = colnames(m))
#m.df$wts <- as.vector(t(wts))
Imat <- matrix(0, nrow = 100, ncol = 10)
colnames(Imat) <- colnames(m)
rownames(Imat) <- with(m.df,
paste0(as.character(Source), "->", as.character(Target)))
for(ix in 1:100){
Imat[ix, m.df[ix, "Source"]] <- -1
Imat[ix, m.df[ix, "Target"]] <- 1
}
Imat <- Imat[complete.cases(m.df), ]
m.df <- m.df[complete.cases(m.df), ]
Dorsal <- c("LM", "P", "LI", "POR")
Ventral <- c("AL", "RL", "A", "PM", "AM")
m.df$Stream <- with(m.df,
ifelse(Source %in% Dorsal, "D",
ifelse(Source %in% Ventral, "V", "V1")))
```



# Beta regression

## p.24: beta regression - print model results 

```{r}
library(betareg)
br1 <- betareg(ODr ~ Imat[, -1] + 0, m.df)
summary(br1)
```


## p.25: beta regression - print 95% confidence intervals
```{r}
br.CI <- 1.96 * coef(summary(br1))$mean[, 2]
( br.ci <- rbind(V1 = 0, 
      cbind(est = coef(summary(br1))$mean[, 1],
            lwr = coef(summary(br1))$mean[, 1] - br.CI, 
            upr = coef(summary(br1))$mean[, 1] + br.CI))
)
```
## p.26: beta regresion - plot areas by hierarchial value with error bars

```{r, fig.width = 6, fig.height = 7}
areas_plotting_order <- c('POR','LI','P','LM','V1','RL','AL','A','PM','AM')
rownames(br.ci) <- rownames(m)
Brhiervals <- c(0, coef(br1)[-10])
class(Brhiervals) <- "hier"
names(Brhiervals) <- rownames(m)
plot(seq_along(Brhiervals), Brhiervals[areas_plotting_order], type = "l",
     ylim = c(-0.1, 1.5))
segments(seq_along(Brhiervals), br.ci[areas_plotting_order, "lwr"], 
         seq_along(Brhiervals), br.ci[areas_plotting_order, "upr"])
text(Brhiervals[areas_plotting_order], labels =
       names(Brhiervals[areas_plotting_order]))
```

## p.26: beta regression 

plot areas' hierarchical values fitted to a scale ranging from 0-10, split into dorsal and ventral streams

```{r}
plot.hier <- function(hier, ...){
  h <- 10 * hier/max(hier)
  Dorsal <- c("LM", "P", "LI", "POR")
  Ventral <- c("AL", "RL", "A", "PM", "AM")
  dstr <- sort(h[names(h) %in% Dorsal])
  vstr <- sort(h[names(h) %in% Ventral])
  plot(c(-1, 1), c(0, 1.1 * max(h)), type = "n",
       axes = FALSE, xlab = "", ylab = "Hierarchical Level")
  axis(2, at = seq(0, 10, 2))
  lines(c(0, rep(-0.5, length(Dorsal))), c(0.05, dstr), type = "c")
  lines(c(0, rep(0.5, length(Ventral))), c(0.05, vstr), type = "c")
  text(0, 0.05, "V1", ...)
  text(rep(-0.5, length(Dorsal)), dstr, names(dstr), ...)
  text(rep(0.5, length(Ventral)), vstr, names(vstr), ...)
}
```

```{r}
plot(Brhiervals, cex = 0.9)
```

## p.27: beta regression 

- print 95% confidence intervals for hierarchical values fitted to a scale ranging from 0-10

```{r}
bSc.fac <- 9/max(coef(summary(br1))$mean[, 1])
bSc.int <- 1.96 * coef(summary(br1))$mean[, 2] * bSc.fac
(bscLhier.ci <- rbind(V1 = 0, 
                     cbind(lwr = bSc.fac * coef(summary(br1))$mean[, 1] + 1 - bSc.int, 
                    upr = bSc.fac * coef(summary(br1))$mean[, 1] + 1 + bSc.int
                    ) )
)
```

## p.27: beta regression

plot fitted OD values vs real OD values

```{r}
par(pty = "s")
plot(fitted(br1) ~ m.df$ODr, 
     xlab = "Proportion of Layer 2-4 labelling",
     ylab = "Fitted Proportion of Layer 2-4 labelling",
     xlim = c(0, 1), ylim =  c(0, 1)
    )
abline(0, 1)
rv <- cor(fitted(br1), m.df$ODr)
text(0, 0.9, bquote(r == .(round(rv, 2))), adj = 0)
```

```{r}
par(pty = "s")
plot(qlogis(fitted(br1)) ~ qlogis(m.df$ODr), 
     xlab = "Proportion of Layer 2-4 labelling",
     ylab = "Fitted Proportion of Layer 2-4 labelling",
     xlim = c(-2, 2), ylim =  c(-2, 2)
    )
abline(0, 1)
rv <- cor(qlogis(fitted(br1)), qlogis(m.df$ODr))
text(-1.9, 1, bquote(r == .(round(rv, 2))), adj = 0)
```


oImat is the incidence matrix with the columns reordered by the estimated hierarchical level from the beta regression.

```{r}
br1.cc <- c(V1 = 0, coef(br1)[-10])
names(br1.cc) <- colnames(Imat)
ord <- order(br1.cc)
oImat <- Imat[, ord]
```

```{r}
mAIC <- sapply(10:2, function(ix){
roImat <- cbind(oImat[, -(ix:10)],
if(ix < 10) rowSums(oImat[, ix:10]) else oImat[, 10] )
m <- betareg(formula = ODr ~ roImat[, -1] + 0, data = m.df)
AIC(m)
})
```

The lowest AIC occurs when all levels from AL to POR are combined.

```{r}
plot(2:10, mAIC, axes = FALSE, ylab = "AIC", xlab = "Areal Cut-off", col="magenta", pch=19, lwd=5) 
axis(2, lwd=2, seq(-200, -150, 5))
axis(1, lwd=2, 2:10, colnames(oImat)[-1])
box()
```





